---
# to run use: rmarkdown::render("nhis_pa_hicp_2020_supplement.Rmd")
title: "Supplement 2: Supplemental Analyses"
subtitle: |
  Physical Activity Guideline Adherence and High-Impact Chronic Pain Among
  U.S. Adults: A Cross-Sectional Analysis of the 2020 National Health
  Interview Survey
author: ""
date: ""
knit: (function(input, ...) rmarkdown::render(input, output_dir = here::here("Output")))
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    number_sections: false
    toc: false
geometry: margin=0.85in
fontsize: 10pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{float}
  - \renewcommand{\arraystretch}{1.25}
  - \renewcommand{\tablename}{}
  - \renewcommand{\thetable}{}
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "asis"
)

# NOTE: Uncomment the line below to install required packages if not already installed.
# install.packages(c("tidyverse","survey","broom","janitor","readr","emmeans","splines","forcats","tableone","knitr","kableExtra","here"))

suppressPackageStartupMessages({
  library(tidyverse)
  library(survey)
  library(broom)
  library(janitor)
  library(readr)
  library(emmeans)
  library(splines)
  library(forcats)
  library(tableone)
  library(knitr)
  library(kableExtra)
  library(here)
})

options(survey.lonely.psu = "adjust")

# Helper: format p-values
fmt_p <- function(p) {
  p <- as.numeric(p)
  ifelse(p < 0.001, "< 0.001", sprintf("%.3f", p))
}
```

```{r load-and-recode}
# ── Load & recode (identical to main script) ─────────────────────────────────
data_path <- here("adult20.csv")
d <- read_csv(data_path, show_col_types = FALSE) |> clean_names()

# Outcome: HICP
d <- d |>
  mutate(
    chronic_pain = case_when(
      paifrq3m_a %in% c(3, 4) ~ 1L,
      paifrq3m_a %in% c(1, 2) ~ 0L,
      TRUE ~ NA_integer_
    ),
    hicp = case_when(
      chronic_pain == 1L & paiwklm3m_a %in% c(3, 4) ~ 1L,
      chronic_pain == 0L ~ 0L,
      chronic_pain == 1L & paiwklm3m_a %in% c(1, 2) ~ 0L,
      TRUE ~ NA_integer_
    )
  )

# Save raw frequency codes BEFORE recoding (for sensitivity analysis)
d <- d |>
  mutate(
    modfreqw_raw = as.numeric(modfreqw_a),
    vigfreqw_raw = as.numeric(vigfreqw_a),
    strfreqw_raw = as.numeric(strfreqw_a)
  )

# Exposure: PA
recode_freq <- function(x) {
  x <- as.numeric(x)
  case_when(
    x %in% c(94, 95)         ~ 0,
    x %in% c(96, 97, 98, 99) ~ NA_real_,
    TRUE                      ~ x
  )
}

recode_min <- function(x) {
  x <- as.numeric(x)
  case_when(
    x >= 996  ~ NA_real_,
    !is.na(x) ~ x,
    TRUE       ~ NA_real_
  )
}

d <- d |>
  mutate(
    modfreqw_a = recode_freq(modfreqw_a),
    vigfreqw_a = recode_freq(vigfreqw_a),
    strfreqw_a = recode_freq(strfreqw_a),
    modmin_a   = recode_min(modmin_a),
    vigmin_a   = recode_min(vigmin_a),
    modmin_a   = ifelse(!is.na(modfreqw_a) & modfreqw_a == 0 &
                          is.na(modmin_a), 0, modmin_a),
    vigmin_a   = ifelse(!is.na(vigfreqw_a) & vigfreqw_a == 0 &
                          is.na(vigmin_a), 0, vigmin_a),
    mod_min_wk = modfreqw_a * modmin_a,
    vig_min_wk = vigfreqw_a * vigmin_a,
    eq_min_wk  = mod_min_wk + 2 * vig_min_wk,
    meet_aerobic = ifelse(!is.na(eq_min_wk) & eq_min_wk >= 150, 1L,
                          ifelse(!is.na(eq_min_wk), 0L, NA)),
    meet_strength = ifelse(!is.na(strfreqw_a) & strfreqw_a >= 2, 1L,
                           ifelse(!is.na(strfreqw_a), 0L, NA)),
    pa_cat = case_when(
      !is.na(eq_min_wk) & eq_min_wk == 0 ~ "No PA",
      !is.na(meet_aerobic) & meet_aerobic == 1 &
        !is.na(meet_strength) & meet_strength == 1 ~
        "Meets aerobic + strength",
      !is.na(meet_aerobic) & meet_aerobic == 1 &
        !is.na(meet_strength) & meet_strength == 0 ~
        "Meets aerobic only",
      !is.na(eq_min_wk) & eq_min_wk > 0 ~
        "Some PA but below guidelines",
      TRUE ~ NA_character_
    ),
    pa_cat = factor(pa_cat, levels = c(
      "No PA",
      "Some PA but below guidelines",
      "Meets aerobic only",
      "Meets aerobic + strength"
    ))
  )

# Covariates — base
d <- d |>
  mutate(
    sex_a = factor(sex_a, levels = c(1, 2), labels = c("Male", "Female")),
    is_hisp = case_when(
      hisp_a == 1 ~ 1L, hisp_a == 2 ~ 0L, TRUE ~ NA_integer_
    ),
    race_eth = case_when(
      is_hisp == 1                     ~ "Hispanic",
      is_hisp == 0 & raceallp_a == 1   ~ "NH White",
      is_hisp == 0 & raceallp_a == 2   ~ "NH Black",
      is_hisp == 0 & raceallp_a == 4   ~ "NH Asian",
      is_hisp == 0                     ~ "NH Other/Multiracial",
      TRUE ~ NA_character_
    ) |> factor(levels = c(
      "NH White", "NH Black", "NH Asian", "Hispanic",
      "NH Other/Multiracial"
    ))
  )

# Covariates — expanded (Model B)
d <- d |>
  mutate(
    educ_cat = case_when(
      educ_a %in% c(0, 1, 2, 3) ~ "Less than HS",
      educ_a == 4                ~ "HS diploma/GED",
      educ_a %in% c(5, 6, 7, 8) ~ "Some college/Associate",
      educ_a %in% c(9, 10)      ~ "Bachelor or higher",
      TRUE ~ NA_character_
    ) |> factor(levels = c(
      "Less than HS", "HS diploma/GED",
      "Some college/Associate", "Bachelor or higher"
    )),
    bmi_cat = case_when(
      bmicat_a == 1 ~ "Underweight",
      bmicat_a == 2 ~ "Normal weight",
      bmicat_a == 3 ~ "Overweight",
      bmicat_a == 4 ~ "Obese",
      TRUE ~ NA_character_
    ) |> factor(levels = c(
      "Normal weight", "Underweight", "Overweight", "Obese"
    )),
    arthritis = case_when(
      arthev_a == 1 ~ 1L, arthev_a == 2 ~ 0L, TRUE ~ NA_integer_
    ),
    dep_freq = case_when(
      depfreq_a %in% c(1, 2) ~ "Daily/Weekly",
      depfreq_a == 3          ~ "Monthly",
      depfreq_a == 4          ~ "A few times a year",
      depfreq_a == 5          ~ "Never",
      TRUE ~ NA_character_
    ) |> factor(levels = c(
      "Never", "A few times a year", "Monthly", "Daily/Weekly"
    )),
    anx_freq = case_when(
      anxfreq_a %in% c(1, 2) ~ "Daily/Weekly",
      anxfreq_a == 3          ~ "Monthly",
      anxfreq_a == 4          ~ "A few times a year",
      anxfreq_a == 5          ~ "Never",
      TRUE ~ NA_character_
    ) |> factor(levels = c(
      "Never", "A few times a year", "Monthly", "Daily/Weekly"
    )),
    income_cat = case_when(
      ratcat_a %in% c(1, 2, 3, 4, 5) ~ "<2.00",
      ratcat_a %in% c(6, 7, 8, 9)    ~ "2.00-3.99",
      ratcat_a %in% c(10, 11)        ~ "4.00-4.99",
      ratcat_a %in% c(12, 13, 14)    ~ "5.00+",
      TRUE ~ NA_character_
    ) |> factor(levels = c("<2.00", "2.00-3.99", "4.00-4.99", "5.00+")),
    has_code95 = (modfreqw_raw == 95 | vigfreqw_raw == 95 |
                    strfreqw_raw == 95),
    has_code95 = ifelse(is.na(has_code95), FALSE, has_code95)
  )
```

```{r analytic-samples}
# ── Analytic samples ──────────────────────────────────────────────────────────
# Base complete-case (Model A)
d_cc <- d |>
  filter(
    !is.na(wtfa_a), !is.na(pstrat), !is.na(ppsu),
    !is.na(hicp), !is.na(pa_cat),
    !is.na(agep_a), !is.na(sex_a), !is.na(race_eth)
  ) |>
  mutate(
    hicp1    = as.numeric(hicp == 1),
    pa_cat   = fct_drop(pa_cat),
    sex_a    = fct_drop(sex_a),
    race_eth = fct_drop(race_eth),
    pa_score = case_when(
      pa_cat == "No PA" ~ 0,
      pa_cat == "Some PA but below guidelines" ~ 1,
      pa_cat == "Meets aerobic only" ~ 2,
      pa_cat == "Meets aerobic + strength" ~ 3
    ),
    eq_capped = pmin(eq_min_wk, 1500),
    str_days  = strfreqw_a,
    age_grp   = case_when(
      agep_a < 40               ~ "18-39",
      agep_a >= 40 & agep_a < 65 ~ "40-64",
      agep_a >= 65              ~ "65+"
    ) |> factor(levels = c("18-39", "40-64", "65+"))
  )

# Expanded complete-case (Model B)
d_cc_exp <- d_cc |>
  filter(
    !is.na(educ_cat), !is.na(bmi_cat), !is.na(arthritis),
    !is.na(dep_freq), !is.na(anx_freq), !is.na(income_cat)
  )

# Sensitivity: exclude code 95
d_cc_no95 <- d_cc |> filter(!has_code95)

n_base    <- nrow(d_cc)
n_exp     <- nrow(d_cc_exp)
n_no95    <- nrow(d_cc_no95)
n_removed <- sum(d_cc$has_code95)
n_excl    <- nrow(d) - n_base
```

```{r survey-designs}
# ── Survey designs ────────────────────────────────────────────────────────────
des_base <- svydesign(
  ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
  nest = TRUE, data = d_cc
)
des_base <- update(des_base, pa_score = d_cc$pa_score,
                   age_grp = d_cc$age_grp)

des_exp <- svydesign(
  ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
  nest = TRUE, data = d_cc_exp
)
des_exp <- update(des_exp, pa_score = d_cc_exp$pa_score)

des_no95 <- svydesign(
  ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
  nest = TRUE, data = d_cc_no95
)
des_no95 <- update(des_no95, pa_score = d_cc_no95$pa_score)
```

## S1. Missingness Comparison: Included vs. Excluded Respondents

Table S1 compares demographic characteristics and HICP status (where
available) between respondents included in the analytic sample
(n = `r format(n_base, big.mark = ",")`) and those excluded due to
missing data (n = `r format(n_excl, big.mark = ",")`). Excluded
respondents were older, more likely to be female, and had substantially
higher HICP prevalence (13.2% vs. 5.1%), suggesting that complete-case
analysis may underestimate population HICP prevalence.

```{r s1-missingness}
d_full <- d |>
  filter(!is.na(wtfa_a), !is.na(pstrat), !is.na(ppsu)) |>
  mutate(
    in_analytic = !is.na(hicp) & !is.na(pa_cat) &
                  !is.na(agep_a) & !is.na(sex_a) & !is.na(race_eth),
    inclusion   = ifelse(in_analytic, "Included", "Excluded"),
    hicp_avail  = ifelse(!is.na(hicp), as.character(hicp), NA_character_)
  )

tab_miss <- CreateTableOne(
  vars       = c("agep_a", "sex_a", "race_eth", "hicp_avail"),
  strata     = "inclusion",
  data       = d_full,
  factorVars = c("sex_a", "race_eth", "hicp_avail"),
  includeNA  = FALSE,
  test       = TRUE,
  addOverall = FALSE
)

mat_miss <- print(tab_miss, showAllLevels = TRUE, quote = FALSE,
                  noSpaces = TRUE, printToggle = FALSE)
df_miss  <- data.frame(
  Characteristic = rownames(mat_miss), mat_miss,
  row.names = NULL, check.names = FALSE
)

df_miss$Characteristic <- df_miss$Characteristic |>
  str_replace("agep_a \\(mean \\(SD\\)\\)", "Age, mean (SD)") |>
  str_replace("sex_a \\(%\\)", "Sex (%)") |>
  str_replace("race_eth \\(%\\)", "Race/Ethnicity (%)") |>
  str_replace("hicp_avail \\(%\\)", "HICP status (%)")

if ("p" %in% colnames(df_miss)) {
  df_miss$p <- sapply(df_miss$p, function(x) {
    if (is.na(x) || x == "") return("")
    val <- suppressWarnings(as.numeric(x))
    if (is.na(val)) {
      if (grepl("^<", x)) return(x)
      return(x)
    }
    fmt_p(val)
  })
}

kable(
  df_miss,
  caption = "Table S1. Unweighted comparison of included vs. excluded respondents.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )
```

*Note:* HICP status available for 726 of `r format(n_excl, big.mark = ",")`
excluded respondents; the remaining lacked classifiable pain or
activity-limitation data. P-values from chi-squared (categorical)
or *t*-test (continuous).

\newpage

## S2. Confounding Assessment: Model A vs. Model B

To assess potential confounding by health-related and socioeconomic
factors, we estimated adjusted HICP prevalences under two models on the
same expanded complete-case sample. Model A adjusted for age, sex, and
race/ethnicity only (total association). Model B additionally adjusted
for education, BMI category, arthritis history, depression frequency,
anxiety frequency, and income-to-poverty ratio (conservative association
after additional adjustment). We note that several Model B covariates
(BMI, arthritis, depression, anxiety) may lie on the causal pathway
between PA and HICP; thus, Model B may represent overadjustment and
should be interpreted as the lower bound of the true association.

```{r s2-confounding}
emm_options(df = survey::degf(des_exp))

# Model A: demographics only (on expanded sample for comparability)
fit_A <- svyglm(hicp1 ~ pa_cat + agep_a + sex_a + race_eth,
                design = des_exp, family = gaussian())
emm_A  <- summary(emmeans(fit_A, ~ pa_cat))

trend_A <- svyglm(hicp1 ~ pa_score + agep_a + sex_a + race_eth,
                  design = des_exp, family = gaussian())
trend_A_test <- survey::regTermTest(trend_A, ~ pa_score)
b_A  <- coef(summary(trend_A))["pa_score", "Estimate"]
se_A <- coef(summary(trend_A))["pa_score", "Std. Error"]

# Model B: fully adjusted
fit_B <- svyglm(
  hicp1 ~ pa_cat + agep_a + sex_a + race_eth +
    educ_cat + bmi_cat + arthritis + dep_freq + anx_freq + income_cat,
  design = des_exp, family = gaussian()
)
emm_B <- summary(emmeans(
  fit_B, ~ pa_cat,
  nuisance = c("educ_cat", "bmi_cat", "dep_freq", "anx_freq", "income_cat")
))

trend_B <- svyglm(
  hicp1 ~ pa_score + agep_a + sex_a + race_eth +
    educ_cat + bmi_cat + arthritis + dep_freq + anx_freq + income_cat,
  design = des_exp, family = gaussian()
)
trend_B_test <- survey::regTermTest(trend_B, ~ pa_score)
b_B  <- coef(summary(trend_B))["pa_score", "Estimate"]
se_B <- coef(summary(trend_B))["pa_score", "Std. Error"]

attenuation_pct <- 100 * (1 - abs(b_B) / abs(b_A))

# Table S2a
s2a <- tibble(
  `PA category`          = as.character(emm_A$pa_cat),
  `Model A: Adj. HICP %` = sprintf("%.1f", 100 * emm_A$emmean),
  `Model A: 95% CI`      = sprintf("%.1f\u2013%.1f",
                                   100 * emm_A$lower.CL,
                                   100 * emm_A$upper.CL),
  `Model B: Adj. HICP %` = sprintf("%.1f", 100 * emm_B$emmean),
  `Model B: 95% CI`      = sprintf("%.1f\u2013%.1f",
                                   100 * emm_B$lower.CL,
                                   100 * emm_B$upper.CL)
)

kable(
  s2a,
  caption = "Table S2a. Adjusted HICP prevalence (\\%) by PA category, Models A and B.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 9
  )
```

```{r s2b-trend}
s2b <- tibble(
  Model = c("A (demographics)", "B (fully adjusted)"),
  `Per-category change (pp)` = sprintf("%.2f", c(100 * b_A, 100 * b_B)),
  `95% CI` = sprintf(
    "%.2f to %.2f",
    c(100 * (b_A - 1.96 * se_A), 100 * (b_B - 1.96 * se_B)),
    c(100 * (b_A + 1.96 * se_A), 100 * (b_B + 1.96 * se_B))
  ),
  `Trend p-value` = c(fmt_p(trend_A_test$p), fmt_p(trend_B_test$p))
)

kable(
  s2b,
  caption = "Table S2b. Linear trend in HICP across PA categories, Models A and B.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 9
  )
```

*Note:* Model A: age + sex + race/ethnicity. Model B: Model A +
education + BMI category + arthritis + depression frequency + anxiety
frequency + income-to-poverty ratio. The per-category change attenuated
by approximately `r round(attenuation_pct)`% from Model A to Model B,
but the linear trend remained highly significant (*p* < 0.001).

\newpage

## S3. Sensitivity Analysis: Excluding Code 95 Respondents

NHIS frequency-per-week recode code 95 denotes respondents who reported
being "unable to do" a given activity type. In the primary analysis,
these respondents were assigned zero frequency (equivalent to code 94,
"never"). Because inability to exercise may itself be caused by pain,
this coding could inflate the observed PA--HICP association. To test
this, we re-estimated the primary models after excluding all respondents
with code 95 on any PA frequency variable (n = `r n_removed` removed;
analytic n = `r format(n_no95, big.mark = ",")`).

```{r s3-sensitivity}
emm_options(df = survey::degf(des_no95))

fit_no95 <- svyglm(hicp1 ~ pa_cat + agep_a + sex_a + race_eth,
                   design = des_no95, family = gaussian())
emm_no95 <- summary(emmeans(fit_no95, ~ pa_cat))

trend_no95 <- svyglm(hicp1 ~ pa_score + agep_a + sex_a + race_eth,
                     design = des_no95, family = gaussian())
trend_no95_test <- survey::regTermTest(trend_no95, ~ pa_score)
b_no95  <- coef(summary(trend_no95))["pa_score", "Estimate"]
se_no95 <- coef(summary(trend_no95))["pa_score", "Std. Error"]

s3a <- tibble(
  `PA category`           = as.character(emm_no95$pa_cat),
  `Adj. HICP % (no 95)`   = sprintf("%.1f", 100 * emm_no95$emmean),
  `95% CI`                = sprintf("%.1f\u2013%.1f",
                                    100 * emm_no95$lower.CL,
                                    100 * emm_no95$upper.CL)
)

kable(
  s3a,
  caption = "Table S3a. Adjusted HICP prevalence excluding code 95 respondents.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

```{r s3b-trend}
s3b <- tibble(
  Analysis = "Excluding code 95",
  `Per-category change (pp)` = sprintf("%.2f", 100 * b_no95),
  `95% CI` = sprintf("%.2f to %.2f",
                     100 * (b_no95 - 1.96 * se_no95),
                     100 * (b_no95 + 1.96 * se_no95)),
  `Trend p-value` = fmt_p(trend_no95_test$p),
  `n removed`     = n_removed,
  `Analytic n`    = format(n_no95, big.mark = ",")
)

kable(
  s3b,
  caption = "Table S3b. Trend test excluding code 95 respondents.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size = 9)
```

*Note:* Results are virtually identical to the primary analysis
(per-category change `r sprintf("%.2f", 100 * b_no95)` pp vs.
&minus;1.82 pp in primary; *p* < 0.001 in both). Excluding code 95
respondents does not materially affect the findings.

\newpage

## S4. Subgroup Analyses: Effect Modification by Age, Sex, and Race/Ethnicity

We tested whether the PA--HICP association varied across demographic
subgroups by fitting interaction terms (PA ordinal score $\times$
subgroup) in survey-weighted linear probability models. Stratified
per-category slopes are reported alongside formal interaction *p*-values.

```{r s4-subgroup}
# ── Age group interaction ────────────────────────────────────────────────────
fit_age_int <- svyglm(hicp1 ~ pa_score * age_grp + sex_a + race_eth,
                      design = des_base, family = gaussian())
age_int_test <- survey::regTermTest(fit_age_int, ~ pa_score:age_grp)

age_slopes <- tibble()
for (ag in levels(d_cc$age_grp)) {
  sub <- d_cc |> filter(age_grp == ag)
  des_sub <- svydesign(ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
                       nest = TRUE, data = sub)
  des_sub <- update(des_sub, pa_score = sub$pa_score)
  fit_sub <- svyglm(hicp1 ~ pa_score + sex_a + race_eth,
                    design = des_sub, family = gaussian())
  b  <- coef(summary(fit_sub))["pa_score", "Estimate"]
  se <- coef(summary(fit_sub))["pa_score", "Std. Error"]
  p  <- coef(summary(fit_sub))["pa_score", "Pr(>|t|)"]
  age_slopes <- bind_rows(age_slopes, tibble(
    Subgroup = ag,
    `Per-category change (pp)` = sprintf("%.2f", 100 * b),
    `95% CI` = sprintf("%.2f to %.2f",
                       100 * (b - 1.96 * se), 100 * (b + 1.96 * se)),
    `p-value` = fmt_p(p)
  ))
}

# ── Sex interaction ──────────────────────────────────────────────────────────
fit_sex_int <- svyglm(hicp1 ~ pa_score * sex_a + agep_a + race_eth,
                      design = des_base, family = gaussian())
sex_int_test <- survey::regTermTest(fit_sex_int, ~ pa_score:sex_a)

sex_slopes <- tibble()
for (sx in levels(d_cc$sex_a)) {
  sub <- d_cc |> filter(sex_a == sx)
  des_sub <- svydesign(ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
                       nest = TRUE, data = sub)
  des_sub <- update(des_sub, pa_score = sub$pa_score)
  fit_sub <- svyglm(hicp1 ~ pa_score + agep_a + race_eth,
                    design = des_sub, family = gaussian())
  b  <- coef(summary(fit_sub))["pa_score", "Estimate"]
  se <- coef(summary(fit_sub))["pa_score", "Std. Error"]
  p  <- coef(summary(fit_sub))["pa_score", "Pr(>|t|)"]
  sex_slopes <- bind_rows(sex_slopes, tibble(
    Subgroup = sx,
    `Per-category change (pp)` = sprintf("%.2f", 100 * b),
    `95% CI` = sprintf("%.2f to %.2f",
                       100 * (b - 1.96 * se), 100 * (b + 1.96 * se)),
    `p-value` = fmt_p(p)
  ))
}

# ── Race/ethnicity interaction ───────────────────────────────────────────────
fit_race_int <- svyglm(hicp1 ~ pa_score * race_eth + agep_a + sex_a,
                       design = des_base, family = gaussian())
race_int_test <- survey::regTermTest(fit_race_int, ~ pa_score:race_eth)

race_slopes <- tibble()
for (rc in levels(d_cc$race_eth)) {
  sub <- d_cc |> filter(race_eth == rc)
  if (nrow(sub) < 100) next
  des_sub <- svydesign(ids = ~ppsu, strata = ~pstrat, weights = ~wtfa_a,
                       nest = TRUE, data = sub)
  des_sub <- update(des_sub, pa_score = sub$pa_score)
  fit_sub <- svyglm(hicp1 ~ pa_score + agep_a + sex_a,
                    design = des_sub, family = gaussian())
  b  <- coef(summary(fit_sub))["pa_score", "Estimate"]
  se <- coef(summary(fit_sub))["pa_score", "Std. Error"]
  p  <- coef(summary(fit_sub))["pa_score", "Pr(>|t|)"]
  race_slopes <- bind_rows(race_slopes, tibble(
    Subgroup = rc,
    `Per-category change (pp)` = sprintf("%.2f", 100 * b),
    `95% CI` = sprintf("%.2f to %.2f",
                       100 * (b - 1.96 * se), 100 * (b + 1.96 * se)),
    `p-value` = fmt_p(p)
  ))
}

# Combine
s4_slopes <- bind_rows(
  age_slopes  |> mutate(Modifier = "Age group", .before = 1),
  sex_slopes  |> mutate(Modifier = "Sex", .before = 1),
  race_slopes |> mutate(Modifier = "Race/Ethnicity", .before = 1)
)

kable(
  s4_slopes,
  caption = "Table S4a. Stratified per-category change in HICP (pp) by subgroup.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size = 9)
```

```{r s4b-interactions}
s4_int <- tibble(
  Modifier        = c("Age group", "Sex", "Race/Ethnicity"),
  `Interaction p` = c(
    fmt_p(age_int_test$p),
    fmt_p(sex_int_test$p),
    fmt_p(race_int_test$p)
  )
)

kable(
  s4_int,
  caption = "Table S4b. Interaction test p-values.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

*Note:* All three interaction terms were statistically significant. The
PA--HICP gradient was strongest among adults aged 40+ and among women.
Among racial/ethnic groups, NH White and NH Asian adults showed the
steepest gradients, though the NH Asian estimate had wide confidence
intervals due to smaller sample size.

\newpage

## S5. Linear Probability Model: Predicted Probability Bounds

A known limitation of the linear probability model (LPM) is that
predicted probabilities can fall outside [0, 1]. We examined the
distribution of fitted values from the primary LPM.

```{r s5-bounds}
fit_lpm_base <- svyglm(hicp1 ~ pa_cat + agep_a + sex_a + race_eth,
                       design = des_base, family = gaussian())
pred_vals <- as.numeric(predict(fit_lpm_base, type = "response"))

s5 <- tibble(
  Metric = c(
    "Min predicted probability",
    "Max predicted probability",
    "N predictions < 0",
    "N predictions > 1",
    "Proportion in [0,1]"
  ),
  Value = c(
    sprintf("%.4f", min(pred_vals, na.rm = TRUE)),
    sprintf("%.4f", max(pred_vals, na.rm = TRUE)),
    as.character(sum(pred_vals < 0, na.rm = TRUE)),
    as.character(sum(pred_vals > 1, na.rm = TRUE)),
    sprintf("%.4f", mean(pred_vals >= 0 & pred_vals <= 1, na.rm = TRUE))
  )
)

n_below_zero <- sum(pred_vals < 0, na.rm = TRUE)
pct_below    <- sprintf("%.1f", 100 * n_below_zero / length(pred_vals))
min_pred     <- sprintf("%.3f", min(pred_vals, na.rm = TRUE))

kable(
  s5,
  caption = "Table S5. LPM predicted probability diagnostics.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

*Note:* `r format(n_below_zero, big.mark = ",")` observations
(`r pct_below`%) had predicted probabilities slightly below zero
(minimum `r min_pred`). No predictions exceeded 1.0. These out-of-range
predictions are substantively negligible and are typical for LPMs
applied to low-prevalence outcomes. The modified Poisson model, which
constrains predictions to the positive real line, serves as a
complementary robustness check.

## S6. Natural Cubic Spline Knot Placement

The dose--response model used a natural cubic spline via `ns(df = 3)`
from the **splines** package for moderate-equivalent minutes/week
(capped at 1,500). With `df = 3`, `ns()` places two internal knots at
the 33rd and 67th percentiles of the observed data distribution.

```{r s6-knots}
# Extract actual knot locations from ns()
ns_obj   <- ns(d_cc$eq_capped, df = 3)
int_knots <- attr(ns_obj, "knots")
bnd_knots <- attr(ns_obj, "Boundary.knots")

# ns(df = 3) places two internal knots at the 33rd and 67th percentiles
s6 <- tibble(
  `Knot type` = c(
    "Boundary (lower)",
    "Internal (33rd percentile)",
    "Internal (67th percentile)",
    "Boundary (upper)"
  ),
  `Location (moderate-eq. min/wk)` = c(
    format(bnd_knots[1], big.mark = ","),
    format(int_knots[1], big.mark = ","),
    format(int_knots[2], big.mark = ","),
    format(bnd_knots[2], big.mark = ",")
  )
)

kable(
  s6,
  caption = "Table S6. Natural cubic spline knot locations from ns(df = 3).",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

*Note:* Internal knots at `r int_knots[1]` and `r int_knots[2]` min/wk
correspond to the 33rd and 67th percentiles of capped moderate-equivalent
minutes/week among the analytic sample. Boundary knots are placed at the
range extremes (`r bnd_knots[1]` and
`r format(bnd_knots[2], big.mark = ",")` min/wk).

## S7. Justification for Capping Moderate-Equivalent Minutes at 1,500 Min/Week

Self-reported PA data can include implausibly high values. We capped
moderate-equivalent minutes/week at 1,500 to reduce the influence of
extreme observations on the spline-based dose--response model.

```{r s7-cap}
n_above_1500 <- sum(d_cc$eq_min_wk > 1500, na.rm = TRUE)
n_above_1000 <- sum(d_cc$eq_min_wk > 1000, na.rm = TRUE)
n_total_pa   <- sum(!is.na(d_cc$eq_min_wk))
max_val      <- max(d_cc$eq_min_wk, na.rm = TRUE)
p99          <- quantile(d_cc$eq_min_wk, 0.99,  na.rm = TRUE)
p995         <- quantile(d_cc$eq_min_wk, 0.995, na.rm = TRUE)

s7 <- tibble(
  Metric = c(
    "Total with valid moderate-eq. min/wk",
    "Maximum (min/wk)",
    "99th percentile (min/wk)",
    "99.5th percentile (min/wk)",
    "N > 1,000 min/wk",
    "% > 1,000 min/wk",
    "N > 1,500 min/wk",
    "% > 1,500 min/wk"
  ),
  Value = c(
    format(n_total_pa, big.mark = ","),
    format(max_val, big.mark = ","),
    sprintf("%.1f", p99),
    sprintf("%.1f", p995),
    format(n_above_1000, big.mark = ","),
    sprintf("%.2f%%", 100 * n_above_1000 / n_total_pa),
    format(n_above_1500, big.mark = ","),
    sprintf("%.2f%%", 100 * n_above_1500 / n_total_pa)
  )
)

kable(
  s7,
  caption = "Table S7. Distribution of extreme PA values.",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

*Note:* The maximum observed value (`r format(max_val, big.mark = ",")`
min/wk) is implausible, equating to ~`r round(max_val / 60 / 7)` hours/day
of moderate activity. The 99th percentile was
`r sprintf("%.1f", p99)` min/wk. Capping at 1,500 min/wk
(≈21.4 hr/wk, or ~3 hr/day) affected
`r format(n_above_1500, big.mark = ",")` respondents
(`r sprintf("%.2f", 100 * n_above_1500 / n_total_pa)`%) and represents
a conservative threshold that retains high-but-plausible values while
excluding reporting errors.

\newpage

## S8. Prevalence Ratios: Models A and B

Prevalence ratios (PR) were estimated using survey-weighted modified
Poisson regression (quasipoisson family, log link) following Zou (2004).
PRs are a commonly used effect measure for cross-sectional studies with
binary outcomes.

```{r s8-pr}
# ── Model A ──────────────────────────────────────────────────────────────────
fit_pr_A <- svyglm(hicp1 ~ pa_cat + agep_a + sex_a + race_eth,
                   design = des_base, family = quasipoisson(link = "log"))

pr_coefs_A <- coef(summary(fit_pr_A))
pa_rows_A  <- grep("^pa_cat", rownames(pr_coefs_A))

s8a <- tibble(
  `PA category` = c("No PA (ref)",
                     sub("^pa_cat", "", rownames(pr_coefs_A)[pa_rows_A])),
  PR = c("1.00", sprintf("%.2f", exp(pr_coefs_A[pa_rows_A, "Estimate"]))),
  `95% CI` = c("ref", sprintf(
    "%.2f\u2013%.2f",
    exp(pr_coefs_A[pa_rows_A, "Estimate"] -
          1.96 * pr_coefs_A[pa_rows_A, "Std. Error"]),
    exp(pr_coefs_A[pa_rows_A, "Estimate"] +
          1.96 * pr_coefs_A[pa_rows_A, "Std. Error"])
  ))
)

kable(
  s8a,
  caption = "Table S8a. Prevalence ratios for HICP by PA category --- Model A (demographics).",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

```{r s8b-pr}
# ── Model B ──────────────────────────────────────────────────────────────────
fit_pr_B <- svyglm(
  hicp1 ~ pa_cat + agep_a + sex_a + race_eth +
    educ_cat + bmi_cat + arthritis + dep_freq + anx_freq + income_cat,
  design = des_exp, family = quasipoisson(link = "log")
)

pr_coefs_B <- coef(summary(fit_pr_B))
pa_rows_B  <- grep("^pa_cat", rownames(pr_coefs_B))

s8b <- tibble(
  `PA category` = c("No PA (ref)",
                     sub("^pa_cat", "", rownames(pr_coefs_B)[pa_rows_B])),
  PR = c("1.00", sprintf("%.2f", exp(pr_coefs_B[pa_rows_B, "Estimate"]))),
  `95% CI` = c("ref", sprintf(
    "%.2f\u2013%.2f",
    exp(pr_coefs_B[pa_rows_B, "Estimate"] -
          1.96 * pr_coefs_B[pa_rows_B, "Std. Error"]),
    exp(pr_coefs_B[pa_rows_B, "Estimate"] +
          1.96 * pr_coefs_B[pa_rows_B, "Std. Error"])
  ))
)

kable(
  s8b,
  caption = "Table S8b. Prevalence ratios for HICP by PA category --- Model B (fully adjusted).",
  booktabs = TRUE, linesep = ""
) |>
  kable_styling(latex_options = c("HOLD_position"), font_size = 9)
```

*Note:* Model A: age + sex + race/ethnicity. Model B: Model A +
education + BMI category + arthritis + depression frequency + anxiety
frequency + income-to-poverty ratio. In Model A, adults meeting both
aerobic and strength guidelines had 74% lower HICP prevalence compared
to inactive adults (PR = 0.26). In Model B, this attenuation yielded a
50% lower prevalence (PR = 0.50), consistent with partial confounding by
health-related covariates.
